
find statistics about the one2many, many2many, many2one genes.
how much data do we have about them?

```{r}
require(readr)
require(dplyr)
require(ggplot2)
require(plyr)
require(grid)
source("C:/Users/inbal/Google Drive/shen-orr/general_functions.r")
data <- readRDS("C:/Users/inbal/Google Drive/shen-orr/all_comp_files.rds")
```

many2many
load groups that have more than MINDATA datapoints.

```{r}
MINDATA = 40
df40 <- find_many2many_groups(data, MINDATA) #more than 40 datapoints
# df0 <- find_many2many_groups(data, 0) #any amount of datapoints
# write.csv(df40,"/storage/md_shenorr/inbaltz/many2many_groups_with40compfiles.csv")
```

for each group create plot of every couple of orthologs.
NO NEED TO RUN THIS. JUST IF I WANT TO SEE SOME PLOTS...
```{r}
for (group_num in 1:max(df40$group)) {
  title = ""
  cols = ceiling(sqrt(nrow(df40 %>% filter(group == group_num))))
  numPlots = nrow((df40 %>% filter(group == group_num)))
  layout <- matrix(seq(1, cols * ceiling(numPlots / cols)),
                   ncol = cols,
                   nrow = ceiling(numPlots / cols))
  grid.newpage()
  pushViewport(viewport(layout = grid.layout(
    nrow(layout),
    ncol(layout),
    heights = if (nchar(title) > 0) {
      unit(c(0.5, rep(5, nrow(layout) - 1)), "null")
    }
    else {
      unit(c(rep(5, nrow(layout))), "null")
    }
  )))
  
  for (row in 1:nrow((df40 %>% filter(group == group_num)))) {
    temp <- data %>% filter(mouse_entrez_ID ==  (df40 %>% filter(group == group_num))[row, "mouse_entrez_ID"] &
                        human_entrez_ID == (df40 %>% filter(group == group_num))[row, "human_entrez_ID"])
    gg <- ggplot(data = temp, aes(x = MM.Ztest, y = HS.Ztest)) +
      geom_point(
        colour = disease_to_color(temp$disease),
        shape = disease_to_shape(temp$disease),
        size = 3
      ) +
      theme(legend.position = "none") +
      labs(x = paste0((df40 %>% filter(
        group == group_num
      ))[row, "mouse_entrez_ID"], " Ztest mouse"),
      y = paste0((df40 %>% filter(
        group == group_num
      ))[row, "human_entrez_ID"], " Ztest human"))

    # Get the i,j matrix positions of the regions that contain this subplot
    matchidx <- as.data.frame(which(layout == row, arr.ind = TRUE))

    print(gg,
          vp = viewport(
            layout.pos.row = matchidx$row,
            layout.pos.col = matchidx$col
          ))
  }
}
```


this plot is just to copy the legend!!!!
```{r}
diseases = unique(data$disease)
df = data.frame(x = 1:length(diseases),
                y = 1:length(diseases),
                disease = diseases)
ggplot(df, aes(x, y)) +
  geom_point(aes(col = disease, shape = disease), size = 4) +
  scale_color_manual(
    name = "a",
    labels = diseases,
    values = c(
      "Adenocarcinoma" = "black",
      "Cancer" = "black",
      "EcoliInfection" = "black",
      "EpilepsyAutism" = "red",
      "Huntington" = "red",
      "LeptospiralInfection" = "red",
      "Lupus" = "blue",
      "MALTlymphoma" = "blue",
      "OvarianCancer" = "blue",
      "StaphInfection" = "pink",
      "Alpha1Anti" = "pink",
      "BrainTumor" = "pink",
      "Burns" = "green",
      "CysticFibrosis" = "green",
      "DMD" = "green",
      "DownSyndrome" = "yellow",
      "Gaucher" = "yellow",
      "IBD" = "yellow",
      "Injury" = "purple",
      "MyotonicDystrophy" = "purple",
      "RA" = "purple",
      "AtopicDermatitis" = "orange",
      "ProstateCancer" = "orange",
      "ASPS" = "orange"
    )
  ) +
  scale_shape_manual(
    name = "a",
    labels = diseases,
    values = c(
      "Adenocarcinoma" = 15,
      "Cancer" = 16,
      "EcoliInfection" = 8,
      "EpilepsyAutism" = 15,
      "Huntington" = 16,
      "LeptospiralInfection" = 8,
      "Lupus" = 15,
      "MALTlymphoma" = 16,
      "OvarianCancer" = 8,
      "StaphInfection" = 15,
      "Alpha1Anti" = 16,
      "BrainTumor" = 8,
      "Burns" = 15,
      "CysticFibrosis" = 16,
      "DMD" = 8,
      "DownSyndrome" = 15,
      "Gaucher" = 16,
      "IBD" = 8,
      "Injury" = 15,
      "MyotonicDystrophy" = 16,
      "RA" = 8,
      "AtopicDermatitis" = 15,
      "ProstateCancer" = 16,
      "ASPS" = 8
    )
  )
```


plot correlations
```{r}
corrs = data.frame()
for (group_num in 1:max(df40$group)) {
  corrs_temp = c()
  for (row in 1:nrow((df40 %>% filter(group == group_num)))) {
    temp <- data %>% filter(mouse_entrez_ID ==  (df40 %>% filter(group == group_num))[row, "mouse_entrez_ID"] &
                              human_entrez_ID == (df40 %>% filter(group == group_num))[row, "human_entrez_ID"])
    r = round(cor(temp$MM.Ztest, temp$HS.Ztest),1)
    # r = floor((((r/0.5)*100)/20)+6)  #find the correct histogram bin
    
    corrs_temp = c(corrs_temp,r)
  }
  corrs = rbind(corrs, data.frame(group_num=rep(group_num, length(corrs_temp)),corrs_temp))
}
corrs_count = count(corrs, vars = c("corrs_temp", "group_num"))

#sort the heatmap by the max(abs(correlation_valls)).
sort_groups = data.frame(new_group_num=1:122,sorting_arg=0)
for (group in 1:max(corrs_count$group_num)){
  sort_groups[group,2] = max(abs((filter(corrs_count,group_num==group))$corrs_temp))
}
sort_groups["old_group_num"] = order(sort_groups$sorting_arg,decreasing = TRUE)
corrs_count <- merge(corrs_count,sort_groups,by.x = "group_num", by.y = "old_group_num")

ggplot(corrs_count, aes(x=corrs_temp, y=new_group_num , fill=freq)) +
  geom_tile()+
  xlab("pearson correlation coefficient")+ ylab("many2many group number")
count(corrs_count$sorting_arg)
hist(corrs_count$sorting_arg, xlab = "max(abs(pearson_correlation coefficient in group))")
```

here I find how many human genes, have mouse genes that are higly correlated with them.
if there is a human gene that is highly correlated with 2 or more mouse genes, I will still only count it once!
CORR_THRESHOLD defines "highly correlated"
```{r}
highest_corrs = c()
for (human_gene in unique(df40$human_entrez_ID)){
  highest_corr = 0
  for (mouse_gene in unique(filter(df40, human_entrez_ID==human_gene)$mouse_entrez_ID)){
    temp <- data %>% filter(mouse_entrez_ID ==  mouse_gene & human_entrez_ID == human_gene)
    r = abs(round(cor(temp$MM.Ztest, temp$HS.Ztest),1))
    if (r>=highest_corr){ 
      highest_corr = r
    }
  }
  highest_corrs = c(highest_corrs,highest_corr)
}

bar_plot=c()
for (CORR_THRESHOLD in c(0,0.1,0.2,0.3,0.4,0.5,0.6)){
  bar_plot = c(bar_plot,length(highest_corrs[highest_corrs >= CORR_THRESHOLD]))
}
data.frame(c(0,0.1,0.2,0.3,0.4,0.5,0.6),bar_plot)
# barplot(bar_plot,names.arg = c(0,0.1,0.2,0.3,0.4,0.5,0.6), xlab = "correlation threshold", ylab = "#human genes with 'good' mouse representatives")

```

one2many

```{r}
one2many <- data %>% filter(ortholog_type == "1:m")
mouse_geneList = unique(one2many$mouse_entrez_ID)

#some genes are annotated as one2many but because of missing data are in our case one2one. here we filter them out.
for (mouse_gene in mouse_geneList) {
  if (length(unique((
    data %>% filter(mouse_entrez_ID == mouse_gene)
  )$human_entrez_ID)) < 2)
    mouse_geneList = mouse_geneList[mouse_geneList != mouse_gene]
  
}
```


```{r}
library(Hmisc)
corrs=c()  #for histogram of all correlation coefficients
for (mouse_gene in mouse_geneList) {
  human_orthologs = unique((data %>% filter(mouse_entrez_ID == mouse_gene))$human_entrez_ID)
  title = ""
  cols = ceiling(sqrt(length(human_orthologs)))
  numPlots = length(human_orthologs)
  layout <- matrix(seq(1, cols * ceiling(numPlots / cols)),
                   ncol = cols,
                   nrow = ceiling(numPlots / cols))
  grid.newpage()
  pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout), heights = if (nchar(title) > 0) {
      unit(c(0.5, rep(5, nrow(layout) - 1)), "null")
    }
    else {
      unit(c(rep(5, nrow(layout))), "null")
    })))
  for (human_gene in human_orthologs) {
    temp <- data %>% filter(mouse_entrez_ID ==  mouse_gene & human_entrez_ID == human_gene)
    if (length(temp$MM.Ztest)>20) corrs = c(corrs, cor(temp$MM.Ztest, temp$HS.Ztest))
    
    gg = ggplot(data = temp, aes(x = MM.Ztest, y = HS.Ztest)) +
      geom_point(
        colour = disease_to_color(temp$disease),
        shape = disease_to_shape(temp$disease),
        size = 3) +
      labs(x = paste0(mouse_gene, " Ztest mouse"), y = paste0(human_gene, " Ztest human"))

    matchidx <- as.data.frame(which(layout == which(human_gene == human_orthologs), arr.ind = TRUE))
    print(gg, vp = viewport(layout.pos.row = matchidx$row, layout.pos.col = matchidx$col))

  }
}
hist(corrs, xlab="correlation coefficient of a pair of orthologs", main="one -> many")
```

plot correlations in heatmap- every line is a one2many group

```{r}
corrs = data.frame()
for (mouse_gene in unique(one2many$mouse_entrez_ID)) {
  corrs_temp = c()
  for (human_gene in unique(filter(data, mouse_entrez_ID==mouse_gene)$human_entrez_ID)) {
    temp <- data %>% filter(mouse_entrez_ID ==  mouse_gene & human_entrez_ID==human_gene)
    if (nrow(temp) > 10){
    r = round(cor(temp$MM.Ztest, temp$HS.Ztest),1)
    corrs_temp = c(corrs_temp,r)
    }
  }
  corrs = rbind(corrs, data.frame(mouse_gene=rep(mouse_gene, length(corrs_temp)),corrs_temp))
}

corrs_count = count(corrs, vars = c("corrs_temp", "mouse_gene"))

#sort the heatmap by the max(abs(correlation_valls)).
sort_groups = data.frame(mouse_gene=unique(corrs_count$mouse_gene))
sort_groups["sorting_arg"]=0
for (group in unique(corrs_count$mouse_gene)){
  sort_groups[sort_groups$mouse_gene==group,]$sorting_arg=max(abs((filter(corrs_count,mouse_gene==group))$corrs_temp))
}
sort_groups = sort_groups[order(sort_groups$sorting_arg,decreasing = TRUE),]
sort_groups["new_group_num"] = 1:length(unique(corrs_count$mouse_gene))
corrs_count <- merge(corrs_count,sort_groups,by.x = "mouse_gene", by.y = "mouse_gene")

ggplot(corrs_count, aes(x=corrs_temp, y=new_group_num , fill=freq)) +
  geom_tile()+
  xlab("pearson correlation coefficient")+ ylab("one2many group number")

```

here I find how many human genes, have mouse genes that are higly correlated with them.
if there is a human gene that is highly correlated with 2 or more mouse genes, I will still only count it once!
CORR_THRESHOLD defines "highly correlated"
```{r}
highest_corrs = c()
for (human_gene in unique(one2many$human_entrez_ID)){
  highest_corr = 0
  for (mouse_gene in unique(filter(one2many, human_entrez_ID==human_gene)$mouse_entrez_ID)){
    temp <- data %>% filter(mouse_entrez_ID ==  mouse_gene & human_entrez_ID == human_gene)
    if(nrow(temp)>10){
      r = abs(round(cor(temp$MM.Ztest, temp$HS.Ztest),1))
      if (r > highest_corr){ 
        highest_corr = r
      }
    }
  }
  highest_corrs = c(highest_corrs,highest_corr)
}

bar_plot=c()
for (CORR_THRESHOLD in c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9)){
  bar_plot = c(bar_plot,length(highest_corrs[highest_corrs >= CORR_THRESHOLD]))
}
data.frame(c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9),bar_plot)
# barplot(bar_plot,names.arg = c(0,0.1,0.2,0.3,0.4,0.5,0.6), xlab = "correlation threshold", ylab = "#human genes with 'good' mouse representatives")

```
many2one

```{r}
many2one <- data %>% filter(ortholog_type == "n:1")
```

plot correlations in heatmap- every line is a human gene

```{r}
corrs = data.frame()
for (human_gene in unique(many2one$human_entrez_ID)) {
  corrs_temp = c()
  for (mouse_gene in unique(filter(data, human_entrez_ID==human_gene)$mouse_entrez_ID)) {
    temp <- data %>% filter(mouse_entrez_ID ==  mouse_gene & human_entrez_ID==human_gene)
    if (nrow(temp) > 10){
    r = round(cor(temp$MM.Ztest, temp$HS.Ztest),1)
    corrs_temp = c(corrs_temp,r)
    }
  }
  corrs = rbind(corrs, data.frame(human_gene=rep(human_gene, length(corrs_temp)),corrs_temp))
}

corrs_count = count(corrs, vars = c("corrs_temp", "human_gene"))

#sort the heatmap by the max(abs(correlation_valls)).
sort_groups = data.frame(human_gene=unique(corrs_count$human_gene))
sort_groups["sorting_arg"]=0
for (gene in unique(corrs_count$human_gene)){
  sort_groups[sort_groups$human_gene==gene,]$sorting_arg=max(abs((filter(corrs_count,human_gene==gene))$corrs_temp))
}
sort_groups = sort_groups[order(sort_groups$sorting_arg,decreasing = TRUE),]
sort_groups["new_group_num"] = 1:length(unique(corrs_count$human_gene))
corrs_count <- merge(corrs_count,sort_groups,by.x = "human_gene", by.y = "human_gene")

ggplot(corrs_count, aes(x=corrs_temp, y=new_group_num , fill=freq)) +
  geom_tile()+
  xlab("pearson correlation coefficient")+ ylab("human genes")

```

here I find how many human genes, have mouse genes that are higly correlated with them.
if there is a human gene that is highly correlated with 2 or more mouse genes, I will still only count it once!
CORR_THRESHOLD defines "highly correlated"
```{r}
highest_corrs = c()
for (human_gene in unique(many2one$human_entrez_ID)){
  highest_corr = 0
  for (mouse_gene in unique(filter(many2one, human_entrez_ID==human_gene)$mouse_entrez_ID)){
    temp <- data %>% filter(mouse_entrez_ID ==  mouse_gene & human_entrez_ID == human_gene)
    if(nrow(temp)>10){
      r = abs(round(cor(temp$MM.Ztest, temp$HS.Ztest),1))
      if (r > highest_corr){ 
        highest_corr = r
      }
    }
  }
  highest_corrs = c(highest_corrs,highest_corr)
  
}

bar_plot=c()
for (CORR_THRESHOLD in c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9)){
  bar_plot = c(bar_plot,length(highest_corrs[highest_corrs >= CORR_THRESHOLD]))
}
data.frame(c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8,0.9),bar_plot)

```

percent agreement:
lets choose CORR_THRESHOLD=0.3 and see what the percent agreement is.

```{r}
CORR_THRESHOLD=0.3
percent_agreements = c()
for (human_gene in unique(filter(data,ortholog_type!="1:1")$human_entrez_ID)){
  for (mouse_gene in unique(filter(data, human_entrez_ID==human_gene)$mouse_entrez_ID)){
    temp <- data %>% filter(mouse_entrez_ID ==  mouse_gene & human_entrez_ID == human_gene)
    if(nrow(temp)>10){
      r = abs(round(cor(temp$MM.Ztest, temp$HS.Ztest),1))
      if (r > CORR_THRESHOLD){ 
        temp <- filter(temp, ((abs(MM.Ztest)>0.1) | (abs(HS.Ztest)>0.1)))
        percent_agreements = c(percent_agreements,sum((temp$MM.Ztest * temp$HS.Ztest)>0)/length(temp$HS.Ztest))
      }
    }
  }
}

hist(percent_agreements,main = "percent agreements on orthologs with correlation threshold>0.3")

percent_agreements = c()
for (human_gene in unique(filter(data,ortholog_type!="1:1")$human_entrez_ID)){
  for (mouse_gene in unique(filter(data, human_entrez_ID==human_gene)$mouse_entrez_ID)){
    temp <- data %>% filter(mouse_entrez_ID ==  mouse_gene & human_entrez_ID == human_gene)
    if(nrow(temp)>10){
      r = abs(round(cor(temp$MM.Ztest, temp$HS.Ztest),1))
      if (r <= CORR_THRESHOLD){ 
        temp <- filter(temp, ((abs(MM.Ztest)>0.1) | (abs(HS.Ztest)>0.1)))
        percent_agreements = c(percent_agreements,sum((temp$MM.Ztest * temp$HS.Ztest)>0)/length(temp$HS.Ztest))
      }
    }
  }
}

hist(percent_agreements,main = "percent agreements on orthologs with correlation threshold<=0.3")

```














